### 2016-06-14 tex/response.tex

\documentclass{article}

\usepackage{fullpage}
\usepackage{parskip}
\usepackage{hyperref}

\title{Response to review: An open framework for the reproducible study of the
iterated prisoner's dilemma.}
\date{\today}

\begin{document}

\maketitle

We would like to note that the reviewers provided a thorough and high quality
review of our paper. We are grateful for this and thankful for all of their
suggestions as we feel it greatly strengthens our manuscript.

\section{Major comments}

\begin{itemize}
\item \textit{suggested title change}: This has been changed to ``An open
    framework for the reproducible study of the iterated prisoner's dilemma''
\item \textit{Implementation \& Architecture section currently discusses how to
    use library, but should be refactored to expand upon architecture}: The
    various examples have been moved to the Reuse section and an extra paragraph
    explaining the structure has been added.
\item \textit{Side-by-side figures appear too small, and should be replaced with
    full-size (perhaps top-bottom) figures}: Figures have been resized.
\item \textit{Discrepancy between version of software in paper and on Zenodo}:
    This has been addressed.
\item \textit{Dependencies not automatically installed or checked via pip}: This
    has been addressed. Here is a url to the code patch (github pull request)
    that fixed this: \url{https://github.com/Axelrod-Python/Axelrod/pull/542}.
\end{itemize}

\section{Minor comments}

We have addressed almost all of the minor comments, these can be viewed in
detail here:
\url{https://github.com/Axelrod-Python/An-open-reproducible-framework-for-the-study-of-the-iterated-prisoners-dilemma/compare/jors...jors-r1-vk}.

Some of these we comment here for ease of review:

Documentation paragraph in the prob end tournament section has been fixed. It
now reads:

\begin{quote}
    We can also view the length of the matches played by each player. The plot
    shows that the length of each match (for each player) is not the same. The
    median length is 4 which is the expected value with the probability of a
    match ending being 0.5.
\end{quote}

The copyright section of the license has been reworded. A discussion of this can
be viewed here: \url{https://github.com/Axelrod-Python/Axelrod/issues/615}.

The code snippets that were not complete (final line for plotting) have been
fixed. Furthermore, an explanation of how to obtain the full tournament results
has also been included.

The discussion group has been made public. This was an oversight.

Regarding the urls that were not correct in the bibliography: we have removed
these for the present time and will leave the formatting/requirements for these
till the editorial process. Note that the ones we have removed correspond to
well referenced published pieces of work and so removing the urls does not
hinder finding the sources.

One comment we have not addressed is with regards to hypothesis not being a
requirement.  Hypothesis is a requirement for development. Cloudpickle is no
longer a requirement following some substantial rewriting of the internals of
the library (resulting in speed, memory and readability improvements).
TestFixtures is also, no longer a requirement.

Other changes include updating the figures showing the number of strategies and
the latest run of the tournament as well as including another author: due to the
fact that they contributed a new strategy while the paper was under review.

Furthermore, we have made some minor changes to the text. These were all minor
typographical changes aimed at improving the readability of the manuscript.

\end{document}